### 2022-10-06 manuscript-revisions/final-submission-cover-letter.qmd
adjusted naming of supplementary-material

---
title: "Final JCGS Cover Letter"
subtitle: "Eye Fitting Straight Lines in the Modern Era"
author: "Emily A. Robinson, Reka Howard, and Susan VanderPlas"
date: "2022"
format: pdf
---

## manuscript.zip

+ `Eye-Fitting-Straight-Lines-in-the-Modern-Era.pdf`
+ `Eye-Fitting-Straight-Lines-in-the-Modern-Era.Rmd`
+ `Eye-Fitting-Straight-Lines-in-the-Modern-Era.tex`
+ `Eye-Fitting-Straight-Lines-in-the-Modern-Era.pdf`
+ `bibliography.bib`
+ `agsm.bst`
+ `images/ydi-stimuli.png`
+ `figure-latex/eyefitting-example-plot-1.pdf`
+ `figure-latex/eyefitting-gamm-residualplots-1.pdf`
+ `figure-latex/eyefitting-lmer-residualplots-1.pdf`
+ `figure-latex/eyefitting-simplot-1.pdf`
+ `figure-latex/ols-vs-pca-example-1.pdf`

## supplementary-material.zip

+ `README.html` 
+ `eyefitting-model-data.csv`
+ `you-draw-it-eyefitting-analysis.Rmd`