### 2022-10-11 manuscript-revisions/final-submission-cover-letter.qmd

---
title: "Final JCGS Cover Letter"
subtitle: "Eye Fitting Straight Lines in the Modern Era"
author: "Emily A. Robinson, Reka Howard, and Susan VanderPlas"
date: "2022"
format: pdf
---

## manuscript.zip

+ `Eye-Fitting-Straight-Lines-in-the-Modern-Era.pdf` Final pdf version of manuscript.
+ `Eye-Fitting-Straight-Lines-in-the-Modern-Era.Rmd` Final rmarkdown document used to write the manuscript with the proper formatting.
+ `Eye-Fitting-Straight-Lines-in-the-Modern-Era.tex` Final latex file outputted from the rmarkdown document.
+ `bibliography.bib` File containing references in bibtex format.
+ `agsm.bst`
+ `images/ydi-stimuli.png` Image for Figure 1 in the manuscript.
+ `figure-latex/eyefitting-simplot-1.pdf` Image for Figure 2 in the manuscript; originally created using R ggplot2 code within the rmarkdown file.
+ `figure-latex/ols-vs-pca-example-1.pdf` Image for Figure 3 in the manuscript; originally created using R ggplot2 code within the rmarkdown file.
+ `figure-latex/eyefitting-example-plot-1.pdf` Image for Figure 4 in the manuscript; originally created using R ggplot2 code within the rmarkdown file.
+ `figure-latex/eyefitting-lmer-residualplots-1.pdf` Image for Figure 5 in the manuscript; originally created using R ggplot2 code within the rmarkdown file. This image is to be printed in color.
+ `figure-latex/eyefitting-gamm-residualplots-1.pdf` Image for Figure 6 in the manuscript; originally created using R ggplot2 code within the rmarkdown file. This image is to be printed in color.

## supplementary-material.zip

+ `README.html` File containing detailed descriptions of the supplementary material.
+ `eyefitting-model-data.csv` De-identified participant data collected in the study and used for analyses.
+ `you-draw-it-eyefitting-analysis.Rmd` The .Rmd code file used to replicate the analysis in this paper.